### 2022-01-29 reports/manuscript_v2.Rmd
updated

---
title: "A scoping review of rodent trapping studies to understand the prevalence of zoonotic disease in West Africa."
date: "`r Sys.Date()`"
author: 
  - "David Simons, Lauren A. Attfield, Kate E. Jones, Deborah Watson-Jones and Richard Kock"
    # url: https://www.dsimons.org
    # orcid_id: 0000-0001-9655-1656
output: 
  # distill::distill_article:
  #   toc: true
  #   toc_depth: 3
  #   toc_float: true
  word_document:
    toc: true
    toc_depth: 3
  # pdf_document:
  #   toc: true
  #   toc_depth: 3
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here())
  })
always_allow_html: yes
bibliography: citations/manuscript_citations.bib
---
```{r, echo=FALSE, include=FALSE}
source(here::here("scripts", "manuscript_library.R"))
```

# Abstract

Rodents are important reservoirs of zoonotic infectious diseases; 255 species are known hosts of 85 zoonotic pathogens. These reservoir species are globally distributed with West Africa containing multiple known and several further predicted reservoir species. Typically, International Union for Conservation of Nature (IUCN) species distribution maps and Global Biodiversity Information Facility (GBIF) presence maps are used to determine regions at risk for zoonotic spillover events. However, these datasets are biased by incomplete sampling. To investigate the impact of rodent species sampling heterogeneity on zoonotic spillover risk we have systematically reviewed rodent trapping studies to produce a contextually rich dataset that can be used to explore this risk at finer spatial resolution. Here we show that sampling of rodents and their pathogens in the West African region are spatially biased by country and trapping habitat. We found that rodent trapping effort was associated with regional population density and was increased in habitats modified by human activity. We compare rodent species ranges produced by the IUCN with rodent presence data from GBIF and with our enriched data including rodent species absence and trapping effort. Furthermore, we report the spatial bias in the investigation of four important rodent zoonoses, Arenaviridae, *Borellia sp.*, *Bartonella sp.* and *Toxoplasma sp.*. Our results highlight that incorporating sampling bias is important when assessing the risk for zoonotic spillover events from rodents. The synthesis of contextually rich rodent trapping data contributes important information that is lacking in IUCN distribution maps and GBIF species presence data. We anticipate this dataset can support the production of more complete spatial risk assessments of zoonotic spillover events. For example, the inclusion of absence data and trapping effort can help identify regions where data sparsity may hinder unbiased risk prediction. Furthermore, these data identify regions in need of comprehensive study for our four targeted rodent zoonoses.

# Introduction

The potential effect of zoonotic infectious disease outbreaks on human health and societies has been dramatically highlighted by the ongoing SARS-CoV-2 pandemic and recent Ebola virus outbreaks. The number of zoonotic spillover events is projected to increase under intensifying anthropogenic pressure such as; increased human populations [@allen_global_2017], increasing urbanisation [@hassell_urbanization_2017], land use change [@@garcia-pena_land-use_2021] and global climate change [@morse_prediction_2012]. In addition, two taxa - rodents (Rodentia) and bats (Chiroptera) - are proposed to contribute to the greatest number of zoonotic pathogens and the greatest risk of novel pathogen emergence [@han_rodent_2015]. Of 2,220 extant rodent species, 244 (10.7%) are described as being reservoirs of 85 zoonoses [@han_global_2016]. Specifically, West Africa has previously been identified as a region at increased hazard for rodent-borne zoonotic spillover events under different projected scenarios [@garcia-pena_land-use_2021]. Rodents are implicated in the transmission pathways of several human diseases in this region, including, Lassa fever, Schistosomiasis and Leptospirosis [@meerburg_rodent-borne_2009].

Rodent species form diverse assemblages, which provide important and beneficial ecosystem services including pest regulation and seed dispersal [@fischer_ecosystem_2018]. The role of rodent species in zoonotic infectious disease spillover or novel pathogen emergence can be viewed as an ecosystem disservice. Rodents typically demonstrate “fast” life histories [@dobson_fast_2007] due to traits such as early maturation and short gestation, which are also associated with zoonotic reservoirs [@han_rodent_2015; @albery_fast-lived_2021]. Further, these traits are prevalent in species that thrive in human dominated landscapes where they displace species that are less likely to be reservoirs of zoonotic pathogen [@gibb_zoonotic_2020]. The widespread occurrence of reservoir species and their proximity to human activity make the description of rodent species assemblages and host-pathogen interactions vitally important to understanding the risk of zoonotic spillover and novel pathogen emergence [@han_global_2016].

Rodent trapping studies provide a useful method to describe rodent population assemblages and survey for potential zoonoses. Studies have been conducted in West Africa to identify novel potential zoonotic pathogens within rodents [@usaid_one_2021] and to investigate the prevalence of known pathogens within known rodent hosts  to understand the distribution and burden of human disease (e.g., for Lassa fever [@fichet-calvet_diversity_2009] and Schistosomiasis [@catalano_multihost_2020]). These studies provide contextually rich information on when, where and under what conditions rodents were trapped that are typically missing from the global datasets [@bovendorp_optimising_2017]. Despite this, these global datasets are typically used to inform species distribution maps (e.g., IUCN, GBIF) and zoonotic spillover and novel pathogen emergence predictions (e.g., EID2 and GIDEON) [@han_global_2016; @smith_global_2014; @pigott_mapping_2014].

Multiple studies have identified potential geographic hotspots where virus and host species diversity may be expected to be at its greatest, this has been used to predict regional zoonotic spillover and novel pathogen emergence risk [@grace_mapping_2012; @han_global_2016; @allen_global_2017]. However, these spatial distributions could be confounded by bias generated by study design and selection of sampling sites [@plowright_sampling_2019]. For example, systematically increased sampling (e.g. more intensive studies over longer time periods) or over-representation of certain habitats (e.g., peri-urban landscapes) could lead to an apparent association between locations and risk that is driven by these factors rather than an underlying host and virus association [@wille_how_2021; @gibb_mammal_2021]. Conversely some regions may not be sampled adequately and therefore be under-represented in these datasets due to sparse human populations or inaccessible habitats, which may result in these regions being inappropriately reported as at low risk of novel pathogen emergence or zoonotic spillover events.

Here, we synthesise rodent trapping studies performed across West Africa and identify the location and habitat types in which they have been conducted, the microorganisms assessed and the host-pathogen associations that have been reported in order to quantify the potential bias and to identify regions requiring further focussed investigation.

# Methods

## Literature search

We conducted a search in Ovid MEDLINE, Web of Science (Core collection and Zoological Record), JSTOR, BioOne, African Journals Online, Global Health and the pre-print servers, BioRxiv and EcoEvoRxiv for the following terms as keywords, no date limits were set:

  1. Rodent OR Rodent trap*
  2. West Africa (or the individual countries)
  3. 1. AND 2.

We searched other resources including the UN Official Documents System, Open Grey, AGRIS FAO and Google Scholar using combinations of the above terms. Searches were run on 2021-03-01.

We included studies if they met all of the following inclusion criteria; i) reported findings from trapping studies where the target was a small mammal, ii) described the type of trap used or the length of trapping activity or the location of the trapping activity, iii) included trapping activity from at least one West African country (Benin, Burkina Faso, Cape Verde, Côte d'Ivoire, Eastern Sahara, Gambia, Ghana, Guinea, Guinea-Bissau, Liberia, Mali, Mauritania, Niger, Nigeria, Senegal, Sierra Leone and Togo), iv) recorded the genus or species of trapped individuals, v) were published in a peer-reviewed journal or as a pre-print on a digital platform or as a report by a credible organisation. We excluded studies if they met any of the following exclusion criteria: i) reported data that were duplicated from a previously included study, ii) no full text available, iii) not available in English. One reviewer screened titles, abstracts and full texts against the inclusion and exclusion criteria. At each stage, a random subset (10%) was reviewed by a second reviewer.

## Data extraction

We extracted data from eligible studies. Extracted variables included i) study identifiers; ii) study aims; iii) trapping methodology; iv) geolocation data; v) method of speciation; vi) trapping locations and dates; vii) trapped species; viii) number of trap-nights and ix) microorganisms/pathogens of interest. The data extraction tool and the resulting dataset is archived and available in [Supplementary table 1](https://docs.google.com/spreadsheets/d/1rQYjHhk6uk1PoKZZVsgFlmuqWGicU2tTisk9ddfAwTM/edit?usp=sharing).

### Location of rodent trapping studies and habitats studied

We extracted GPS locations for the most precise location presented (i.e., trap, trap-line, study site or study region) by extracting coordinates in the format reported and converting them to decimal degrees. We recorded the habitat classification scheme used by a study (e.g., IUCN Habitat Classification Scheme (Version 3.1)). For studies not using standardised habitat recording, the explicit description of the habitat in which the trap was placed was extracted. For studies reporting multiple habitat types (e.g., rice field, corn field and vegetable garden) for a single trap, trap-line or trapping grid, a higher order classification of habitat type was recorded (e.g., agricultural land).

### Rodent presence, absence, abundance

We mapped genus and species names to the species names used in the Global Biodiversity Information Facility (GBIF) taxonomy  [@gbif_the_global_biodiversity_information_facility_global_2021]. We extracted information on the presence, absence and number of trapped individuals. For studies reporting on all trapped individuals (i.e., not those only reporting on the presence of a specific species of interest), the pseudo-absence of a species reported as present elsewhere in the study was explicitly recorded as an absence at that trap location.

### Pathogen presence and absence

We extracted data on all pathogens assayed in studies investigating rodents for potential zoonoses. The number of rodents tested and the number of positive or negative samples were recorded alongside the type of assay used (e.g., Polymerase Chain Reaction (PCR), Enzyme Linked ImmunoSorbent Assay (ELISA) or viral culture). If studies reported indeterminate results this was noted. Where possible, pathogens were identified to species level. However, where an assay only allowed for attribution to a family of viruses or bacteria, the higher order grouping was used (e.g., Arenaviridae for a PCR using a non-specific arenavirus primer).

## Analysis

### Location and habitats of rodent trapping to investigate potential biases

We summarised the number of studies, the year in which trapping occurred and the country in which they were conducted. We used the GPS coordinates of single trap, trapping grid/line or study site and the number of trap nights to calculate trapping effort (trap night density) within level 2 administrative areas in West Africa. The median trap success rate for all rodents at a defined trap site was calculated separately for trap sites which included built environments and non-built environments due to important differences in trap success between these different environments. The number of rodents trapped at a trap site was then used in combination with these values of trap success to impute the number of trap nights for trap sites where there was no record of the number of trap nights. We summarised the habitat types of trap sites based on information reported in the study; the dictionary used to match habitat types is available in Supplementary 2.

For studies investigating rodent zoonoses we compared the location of trapping sites with SEDAC Global Population Density estimates for 2005, the median year studies were started [@socioeconomic_data_and_applications_center_gridded_2021]. We used a Generalised Additive Model (GAM) incorporating a spatial interaction term to investigate the association between number of trap nights and human population density [@pedersen_hierarchical_2019]. The model structure was specified as:

$$
\begin{equation}
\tag{1}
\text{Trap night density} \sim \text{Tweedie(log(Population density_{2005}) + (Longitude * Latitude))}
\end{equation}
$$

The linear predictor output of this model is the trapping density adjusted for human population density in 2005. We performed sensitivity analysis by removing sites with imputed trapping effort.

We obtained land cover classifications from the European Space Agency Copernicus dataset at 300m^2^ spatial resolution generated using Copernicus Climate Change Service information (2005), then extracted the proportion of land cover classes within all regions of West Africa and all regions in which rodent trapping occurred for investigation of zoonotic diseases. We compared the proportion of possible land cover classes with those of regions where trapping occurred.

Finally, we mapped the presence and absence of rodent species and compared this graphically by overlaying trapping locations to the presence produced from GBIF [@gbif_the_global_biodiversity_information_facility_occurrence_2021] and IUCN [@iucn_iucn_2021] datasets to give a measure of the extent to which each species' range has been sampled, understand the discrepencies between these data sources and to understand the potential impact of biases within these data on subsequently produced models.

### Rodent pathogen associations

We summarised the assays used for microorganisms, the presence and absence of these microorganisms, their host species and the locations from where the samples were obtained. We described the different assays used for each microorganism across included studies. The presence of a microorganism was determined by a report of a positive assay (e.g., PCR, serology, viral culture) in a rodent sample at a defined location and absence of a microorganism was determined by negative results in all rodent samples at a defined location. We described the number of microorganisms tested for in each rodent species and explored the geographic distributions these samples were obtained from. We investigated the association between rodent species and the detection of potential pathogens and reported the proportion of positive and negative tests for each species and pathogen pair.

# Results

## Included studies

We identified 4,282 relevant citations, with 126 rodent trapping studies included. The list of included studies in available in Supplementary 3 or through the interactive web-based application produced to explore this dataset, available [here](https://diddrog11.shinyapps.io/scoping_review_app/). The earliest included studies were published in 1974 with increasing numbers of studies being performed annually since 2005. The mean length of rodent trapping activity was 1.8 years (range 0-23 years). The median time from completion of rodent trapping to publication is 3 years (IQR 2-5 years) (Figure 1.).

![**Year trapping activity begun, ended and the year of publication for included studies.** Each row represents one of the 126 included studies. Green points indicate when the trapping was initiated and blue points indicated the year trapping was concluded. For studies completed within one year the blue point overlies the green. Studies with a transparent grey point did not report the year in which trapping was conducted. The year of publication is indicated by a red point.](../figures/Fig_1_Panel_A.png)

## Location and habitats of rodent trapping studies to investigate potential biases

Rodent trapping took place at 1,331 trap sites with at least one trap site recorded from 14 West African countries. No rodent trapping studies were identified from Gambia, Saint Helena, Ascension and Tristan da Cunha or Togo. The most frequently studied countries were Senegal, Ghana, Guinea, and Nigeria (more than 10 studies each). Fewer than 5 studies were conducted in Liberia, Burkina Faso, and Guinea-Bissau. Thirty-one (25%) studies reported trapping at a single study site, 46 (37%) studies trapped at between two and five study sites, and the remaining 49 studies trapped at between six and 93 study sites. Trap sites were situated in 273 (19.3%) of 1414 level 2 administrative regions in the 14 West African nations. The areas with highest trap night density included the capital cities of Niger (Niamey), Sierra Leone (Freetown), Senegal (Dakar), Mali (Bamako), and Ghana (Accra) and the largest cities of Côte d'Ivoire (Abidjan), and Benin (Cotonou). Outside of these cities, Northern Senegal (Fatique, Thies, Saint-Louis and Kédougou), Southern Guinea (Kindia and Nzerejore), Edo and Osun States in Nigeria and Eastern Sierra Leone had the highest density of trap nights (Figure 2. A).

Greater numbers of trap nights were conducted in areas with higher population densities (*R^2^* = 0.26), adjusting for spatial autocorrelation (Figure 2. B). In particular Mauritania, Northern Senegal, Eastern Sierra Leone, Eastern Guinea and South West Nigeria were trapped at higher rates than would be expected based on their population densities. South West Nigeria, Northern Nigeria, Liberia, Côte d'Ivoire, Ghana, Niger, and Burkina Faso were trapped at lower rates than would be expected. Excluding sites with imputed trap nights reduced the number of trapping sites contributing data. Despite this limitation, Mauritania, Northern Senegal and Sierra Leone remained as regions trapped at higher than expected rates with Nigeria as being trapped at lower than expected rates (Supplementary Figure 1.).

![**Regional trap night density.****A** West Africa mapped to level 2 administrative areas. The colour of the region corresponds to trap night density per 1000 km^2. Purple regions correspond to low trap night density and yellow regions correspond to high trap night density. Colourless regions are level 2 administrative areas with no reported trapping. **B** Map of West Africa coloured by trapping density adjusted for human population density (2005) obtained from our GAM (equation 1). Yellow regions represent areas with higher-than-expected trapping effort for the regional population density and purple regions represent areas trapped at lower-than-expected rates based on their population density. Increased transparency of shading indicates uncertainty in the linear predictor introduced by the smoothing terms on geographical coordinates and population density.](../figures/Figure_2.png)

No studies reported trap habitats with reference to a standardised habitat classification scheme. Extracted habitat types were grouped into 30 categories (see Supplementary Material 2 for the habitat dictionary). At least one habitat was recorded for 17,122 trap sites (95%), with two or more habitats for a single trap site recorded for 4,403 (24%) sites. Single trap sites could span multiple habitat types resulting in 22,202 distinct habitat and trap sites. The most commonly trapped sites were in or around buildings (29%), in areas described as the rodents “natural habitat” (24%), in agricultural areas (13%) (e.g., rice fields, palm plantations), forests (8%) and in the area surrounding buildings (4%) (Figure 3.).

![**Regional land cover of West Africa comparing all regions and trapped regions.** The proportion of the 10 land cover classes reported by the ESA Copernicus dataset (2005) at 300m^2^ resolution are shown for all regions of West Africa (yellow) and for regions in which trapping occurred (purple). Rodent trapping occurred in regions over-representative for cropland, mosaic landscapes and shrubland wile occurring in regions under-representative for bare- and sparse vegetation land cover classes.](../figures/Figure_3.png)

## Rodent presence, absence, abundance

Overall, 4,017 small mammals were trapped (excluding 592 which were trapped outside West African countries). Of these trapped individuals, 2,830 (4%) were identified to order level (Rodentia), 7,283 (10%) were identified to genus level, and the remaining 63,904 (86%) were identified to species level. In studies reporting the number of trap nights the median trap success rate of traps placed in or around buildings was 13% (IQR 6-24%), compared with a median trap success rate of 3% (IQR 1-9%) in other habitats. The majority of the 145 species trapped were Rodentia (112) (Supplementary Material 3), and of these Muridae (82) were the largest family of rodents, followed by; Sciuridae (10), Gliridae and Nesomyidae (both 6), Ctenodactylidae, Anomaluridae, Dipodidae, Hystricidae and Thryonomidae (2 or fewer). The remaining 33 species were from the orders of Soricomorpha (30), Erinaceomorpha (2) and Afrosoricida (1).

*Mastomys sp.* were the most commonly trapped genus (27,079, 37.7%), followed by *Rattus sp.* (11,472, 16.1%), *Mus sp.* (8,624, 12.0%), *Arvicanthis sp.* (5,821, 8.1%) and *Praomys sp.* (5,409, 7.5%). *Mastomys natalensis* were the most commonly trapped species (11,116, 17.4%), followed by, *Rattus rattus* (9,959, 15.6%), *Mastomys erythroleucus* (7,386, 11.6%), *Mus musculus* (6,245, 9.8%), *Arvicanthis niloticus* (5,497, 8.6%), *Mastomys huberti* (4,699, 7.4%) and *Praomys daltoni* (1,854, 2.9%).

We identified locations of the presence or absence points of the seven most commonly trapped species, including true absences and pseudo-absences dependent on the trapping effort at the site. When comparing these with species presence obtained from GBIF and IUCN species range data (Figure 4.) we found that our data were obtained from more geographic locations within a species predicted range than within the GBIF data which were overrepresented from some West African countries, specifically Benin. We observed that our included studies combined with GBIF data suggested that species have greater ranges than were projected from IUCN range maps. This was particularly evident for the invasive species, *Mus musculus* and *Rattus rattus*, which were projected to have more limited distributions within West Africa. Further, we identified species absences were interspersed within presence locations obtained from both trapping studies and GBIF presence locations.

![**Rodent presence, absence and species ranges from this study, GBIF and IUCN.** Each row corresponds to a single rodent species. The left column shows the presence (purple points) and absence (yellow points) of a rodent species from the individual studies included in this review. The centre column shows the presence (green points) of a rodent species obtained from GBIF (September 2021) for records where longitude and latitude have been provided. The right column shows the range of rodent species as proposed by the IUCN (2021) (red shaded area), overlaid with the presence points from both this review and GBIF records.](../figures/Figure_4.png)

## Pathogen presence and absence

Sixty-two studies presented data on microorganisms that infect or are carried by small mammal species in West Africa. Seven studies solely investigated pathogens of rodents, including Hydatigera species (previously Taenia species) and Trichuris species. Zoonotic pathogens and potentially zoonotic microorganisms were investigated in 55 studies. Thirty-two microorganisms were tested for, 8 of which were specified to species level, with the remaining 24 specified at higher taxonomic levels.

Polymerase Chain Reaction (PCR) was used in 32 studies to detect the presence of 22 different species or families of microorganisms from 21,953 rodent samples. The most common microorganisms assayed for using PCR were *Lassa mammarenavirus* or other Arenaviridae (31%), the bacteria *Borrelia sp.* (11.3%) and *Bartonella sp.* (6.5%), followed by Usutu virus (6.4%) and Hantaviridae (5.6%). Molecular tests (antibody or antigen based) were used in 11 studies to detect the presence of 9 different species or families of microorganisms from 11,430 rodent samples. The most common pathogens assessed for with serology were *Lassa mammarenavirus* or other Arenaviridae (78%), *Toxoplasma gondii* (10.6%), *Borrelia sp.* (6.3%) and *Leptospirosis sp.* (2.2%). Histological or direct visualisation assays were used in 8 studies investigating 11 parasitic or bacterial species from 11,229 rodent samples. The most common pathogens assessed for with histology or direct visualisation were *Borrelia sp.* (48.7%), *Schistosoma sp*. (20.4%) and other parasites. Direct culture of *Lassa mammarenavirus* or Leishmania species were reported from 3 studies to detect the presence of these pathogen in 643 rodent samples.

Most (39) studies tested for a single microorganism while 16 studies tested for two or more microorganisms. The most frequently investigated microorganisms were *Lassa mammarenavirus* (17 studies) or members of the Arenaviridae family (10), the spirochete bacteria *Borrelia sp.* (8), *Bartonella sp.* and *Toxoplasma gondii* (4), the remaining 25 microorganisms were reported in three or fewer studies.

The location of trapped rodents that were tested for the four most commonly assayed microorganisms did not encompass the entire range of these hosts. The presence and absence of Arenaviridae, predominantly *Lassa mammarenavirus*, in rodents was investigated in Benin, Côte d'Ivoire, Guinea, Mali, Mauritania, Nigeria, Senegal and Sierra Leone. No studies investigated the prevalence of these pathogens in hosts from Burkina-Faso, Guinea-Bissau, Liberia, Niger, Senegal and Togo. Arenaviridae were detected in rodents all the countries in which it was investigated except Benin and Mauritania (Figure 5A.). The presence and absence of *Borrelia sp.* was investigated in Benin, Mali, Mauritania, Niger and Senegal. *Borrelia sp.* were detected in 8 of 44 species or genera of rodents and shrews assayed in Mali and Senegal (Figure 5B.). The presence and absence of *Bartonella sp.* was investigated in Mali and Nigeria and Senegal with positive samples detected at all but two study sites in Senegal (Figure 5C.). The presence of *Toxoplasma gondii* was investigated in Cape Verde (not shown), Nigeria and Senegal and was detected in 6 of 8 rodent and shrew species assayed (Figure 5D.).

![**Presence and absence of the four most commonly assayed microorganisms.** Arenaviridae (**A**), *Borrelia sp.* (**B**)  *Bartonella sp.* (**C**) and *Toxoplasma gondii* (**D**). The tables to the right of each map highlight the most commonly positive and most commonly assayed rodent species and genera.](../figures/Figure_5_combined.png)

## Host-pathogen associations

Ninety-seven rodent species were investigated for the presence of a microorganism of interest, from 54 studies, in which 42,940 assays were performed on 32,014 individual rodent samples. The rodent species most commonly assayed for these microorganisms were; *Rattus rattus* (n = 2,977) assessed for 24 microorganisms, *Mus musculus* (n = 3,402) for 23, *Mastomys natalensis* (n = 7,189) for 19, and *Mastomys erythroleucus* (n = 3,013) for 19 and *Arvicanthis niloticus* (n = 3,840) for 18. All remaining species were investigated for 10 or fewer microorganisms.

*Lassa mammarenavirus* was detected in 14 species of rodents and Crocidura genera. *Mastomys natalensis* was the most commonly tested species with high rates of positive samples (14%) (Figure 6.). Fifty-four species and genera were tested for arenaviruses, with 37 entirely testing negative. *Bartonella sp.* infections were identified in all 11 species of rodents and shrews that were tested. *Borrelia sp.* infections were identified in 13 species of rodents and shrews. *Mastomys natalensis* were found to have the highest seroprevalence (11%). *Toxoplasma gondii* was detected in 6 species of rodents and shrews, with all individuals from two species *Arvicanthis niloticus* and *Praomys daltoni* testing negative. Finally, 38 species of Rodentia, 3 species of Soricomorpha and 2 species of Erinaceomorpha that were tested for microorganisms had entirely negative results.

![**Number of individuals tested for each microorganism and proportion positive.** For the 27 most commonly assayed rodent species and genera, the percentage of positive assays against each of the 32 microorganisms tested for is shown. The size of a point is proportional to the number of individual animals assayed for that microorganism.](../figures/Figure_6.png)

# Discussion

We identified 124 rodent trapping studies conducted in the West African region since 1965. These studies were conducted in 14 countries, with 54 studies providing information about zoonotic pathogens in trapped rodents. Studies were conducted in all West African countries except Gambia and Togo with between- and within-country variability observed in the number of trap nights conducted. Greater than 70,000 rodents identified to genus or species level were trapped from 112 species of Rodentia. Thirty-two microorganisms were investigated across all studies, and five rodent species were assayed for at least 18 microorganisms. We identify several important limitations to available data, potentially biasing previous zoonotic risk studies; first, in the timing of studies and the locations in which they've been conducted, second, in the locations where a species was recorded as present, and third, in the limited selection of pathogens assayed at these trapping sites.

First, bias is introduced through the absence of a systematic approach to rodent trapping across time and location. The variability in sampling effort made by year and location must be accounted for when attempting to infer the risk of zoonotic spillover events or novel pathogen emergence [@lloyd-smith_nine_2015]. The majority of studies have been conducted since 2000 during which time human population numbers, urbanisation, agricultural- and livestock intensification have progressed dramatically across West Africa [@acheampong_deforestation_2019; @connolly_extended_2021]. These changes alter the availability of habitats for rodent species, potentially increasing the prevalence of commensal species which are more likely to be reservoirs of zoonoses. Thus, accounting for the year in which data were collected and the length of time of the study is important when modelling risk of zoonotic spillover or novel pathogen emergence based on known reservoir species distributions. In addition, several regions, namely North West Senegal, Eastern Sierra Leone, Southern Ghana, Southern Benin, South West and Central Nigeria have been sampled at much higher intensity than elsewhere in West Africa. This sampling heterogeneity will again introduce bias when inferring current or projected risk of zoonotic spillover or novel pathogen emergence [@plowright_sampling_2019].

Second, rodent species were found to have greater species ranges than those reported by IUCN data. Trapping studies were conducted from more regions across West Africa than were available from GBIF records. This increased spatial resolution and the inclusion of presence-absence data presents a more complete representation of the spatial heterogeneity of rodent species' locations within their ranges. Failing to incorporate this heterogeneity in favour of using ranges or presence only data to produce rodent reservoir species' distributions can bias estimations of risk of zoonotic spillover or novel pathogen emergence through neglecting these important confounding factors. For example, the apparent absence of *Mastomys natalensis* at several trapping sites in Nigeria may explain some of the observed spatial variation in human Lassa fever cases that would not be expected based on this species IUCN range or presence only data from GBIF [@redding_geographical_2021].

Third, while 32 potential zoonoses were investigated in at least one study, only four pathogen species or groups were investigated in more than three. Of these four pathogens, studies for *Lassa mammarenavirus* or other Arenaviridae were the most widely performed, both in absolute numbers but also by number of countries and number of study sites within countries. In contrast, *Toxoplasma sp.* were found to be prevalent in rodent hosts within Nigeria and Senegal - despite a substantial known, burden of human disease [@bigna_global_2020] - no studies investigated the prevalence of this pathogen in other West African countries. The limited geographic scope of sampling for these pathogens hampers efforts to understand the risk and human health burden of zoonotic spillover across this diverse region.

Finally, we found that only five of the 112 rodent species trapped had been tested for more than half of the potential pathogens (Figure 6.). There is increasing emphasis on the role of co-infection and multi-species pathogen transmission systems on zoonotic disease spillover risk which depends on comprehensive testing of both known and unknown reservoir species [@sweeny_synzootics_2021]. More comprehensive testing of potential reservoir species across West Africa is required to better understand the role of individual species in disease systems [@douglass_how_2016]. This will also reduce the risk of bias in predictive risk modelling by better characterising host-pathogen associations in species rich environments.

## Limitations

We identified several limitations to our approach. We collated studies from a diverse range of sources but it is possible that rodent surveys that were performed for additional purposes would not be captured in our survey of the scientific and "grey" literature, such as for risk assessments or for  environmental impact assessments for development work or resource extraction. Collated data were limited by data availability from included studies, particularly for measuring trapping effort and identifying habitat types trapped. However, we found that this did not bias our investigation into the association between human population density and trapping effort. Habitat type and human population density were obtained from 2005, the median year in which studies were conducted. Since changes in both human populations and land use in the studied period (1969-2020) have been large across West Africa, this introduces the possibility that these values were not appropriate for the region at the time when the rodent trapping study was conducted.

## Implications

Here we have identified important biases in the locations of studies, the trapping effort for different rodent species and the selection of zoonotic pathogens assayed across the West African region. The use of these data to infer risk of zoonotic spillover or novel disease emergence is likely inappropriate unless these systematic biases are addressed. We hope that this synthesis of current data will encourage further comprehensive ecological studies on rodent species assemblages and their pathogens. We would encourage studies to report uncertainty introduced by these potential biases in model outputs. One priority would be to better understand the true distribution of known rodent reservoirs (particularly *Mastomys natalensis*) across the region with particular emphasis made to the habitat in which they reside and the structure of rodent assemblages.

# Conclusion

In conclusion we present a comprehensive synthesis of rodent trapping studies across West Africa, identifying important biases in the sampling of rodents and their pathogens that limit inference able to be drawn from current evidence. We identify evidence gaps in the current literature and highlight geographic regions, habitat types and host-pathogen associations that would benefit from targeted research effort.

# Acknowledgments
The European Commission nor ECMWF are not responsible for any use that may be made of the Copernicus information or data it contains.

# Supplementary table 1

**Data extraction tool**

# Supplementary table 2

**Habitat dictionary**

# Supplementary table 3

```{r supplementary-3, include=TRUE, echo=FALSE}
supplementary_table <- readRDS(here::here("./data_clean/supplementary_3.rds")) %>%
  dplyr::select(-Author)

set_flextable_defaults(
  font.family = "Calibri",
  font.size = 8,
  font.color = "black",
  theme_fun = "theme_booktabs",
  fonts_ignore = TRUE
)

flextable(supplementary_table) %>%
  colformat_md(j = c("Reference", "Link")) %>%
  colformat_date(j = "Year publication", fmt_date = "%Y") %>%
  autofit()
```

# Supplementary figure 1
![**Regional trap night density. Study sites where trap nights were imputed based on the number of rodents caught were removed from the model. This map of West Africa coloured by trapping density adjusted for human population density (2005) were obtained from our GAM (equation 1) in sensitivity analysis. Yellow regions represent areas with higher-than-expected trapping effort for the regional population density and purple regions represent areas trapped at lower-than-expected rates based on their population density. Increased transparency of shading indicates uncertainty in the linear predictor introduced by the smoothing terms on geographical coordinates and population density.](../figures/Figure_2_sensitivity.png)

# References