### 2022-12-08 reports/archived/test.Rmd
archiving old scripts, manuscript versions and figures

```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      crop = knitr::hook_pdfcrop)
source(here::here("scripts", "manuscript_library.R"))
```

```{r table-1}

read_rds(here("tables", "Table_1_updated.rds")) %>%
  set_caption(caption = "Table 1: Regional")

```

```{r table-1-pixel}

read_rds(here("tables", "Table_1_updated_pixel.rds")) %>%
  set_caption(caption = "Table 1: Pixel")

```